%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% appendix1.tex
%% NOVA thesis document file
%%
%% Chapter with example of appendix with a short dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE appendix1.tex}%

\chapter{Complete Case Studies}
\label{app:appendix}

\section*{Tail Recursive Factorial}

\subsection*{OCaml + GOSPEL}

\begin{gospell}
(*@ 
  function rec factorial (n:int) :int = 
    if n=0 then 1 else n * factorial (n-1)
*)
(*@ 
  requires n >= 0
  variant n
*)

let rec fact_aux n c t =
  if c <= n then fact_aux n (c+1) (t*c) else t
(*@
  r = fact_aux n c t
  requires n >= 0
  requires 0 < c <= n+1
  requires t = factorial (c-1)
  ensures r = factorial n
  variant n+1-c
*)

let fact n = fact_aux n 1 1
(*@ 
  r = fact n
  requires n >= 0 
  ensures r = factorial n
*)
\end{gospell}

\newpage

\subsection*{Generated CakeML}

\begin{cakeml}
fun fact_aux n c t = let val n1 = n in
  let val c1 = c in
  let val t1 = t in
  if c1 <= n1 then (fact_aux n1 (c1 + (1)) (t1 * c1))  else (t1)

fun fact n = let val n1 = n in fact_aux n1 (1) (1)
\end{cakeml}

\subsection*{Corrected CakeML 1st Version}

\begin{cakeml}
fun fact_aux n c t = let val n1 = n in
  let val c1 = c in
  let val t1 = t in
  if c1 <= n1 then (fact_aux n1 (c1 + (1)) (t1 * c1))  else (t1)
  end end end

fun fact n = let val n1 = n in fact_aux n1 (1) (1) end
\end{cakeml}

\subsection*{Corrected CakeML 2nd Version}

\begin{cakeml}
fun fact_aux n c t = 
  if c <= n then (fact_aux n (c + (1)) (t * c))  else (t)

fun fact n = fact_aux n (1) (1)
\end{cakeml}

\newpage

\section*{Exception recursive search}

\subsection*{OCaml + GOSPEL}

\begin{gospell}
let rec search_aux a c n =
  let exception Break of int in try
    if c = Array.length a then raise Not_found 
    else if a.(c) = n then raise (Break c) 
    else search_aux a (c+1) n
  with Break i -> i
(*@
  r = search_aux a c n
  requires 0 <= c <= Array.length a
  requires forall k. 0 <= k < c -> a.(k) <> n
  raises Not_found -> forall k. 0 <= k < Array.length a -> a.(k) <> n
  ensures 0 <= r < Array.length a
  ensures a.(r) = n
  variant Array.length a - c
*)

let linear_search a n = search_aux a 0 n
(*@
  r = linear_search a n
  raises Not_found -> forall k. 0 <= k < Array.length a -> a.(k) <> n
  ensures 0 <= r < Array.length a
  ensures a.(r) = n
*)
\end{gospell}

\subsection*{Generated CakeML}

\begin{cakeml}
fun search_aux a c n = let val a1 = a in
  let val c1 = c in
  let val n1 = n in
  let exception Break of (int) in
  ((if c1 = (a1.length) then (raise Not_found) 
    else (if (get a1 c1) = n1 then (raise (Break n1)) 
          else (search_aux a1 (c1 + (1)) n1)))
  handle   Break i => i)

fun linear_search a n =
  let val a1 = a in let val n1 = n in search_aux a1 (0) n1
\end{cakeml}

\newpage

\subsection*{Corrected CakeML}

\begin{cakeml}

exception Not_found
exception Break int

fun search_aux a c n =
  ((if c = (Array.length a) then (raise Not_found) 
    else (if (Array.sub a c) = n then (raise (Break n)) 
          else (search_aux a (c + (1)) n)))
  handle Break i => i)

fun linear_search a n = search_aux a (0) n

\end{cakeml}

\newpage

\section*{Count Even For A List}

\subsection*{OCaml + GOSPEL}

\begin{gospell}
let count_even l =
  let l1 = List.filter (fun x -> x > 0 && x mod 2 = 0) l in
  let l2 = List.rev l1 in
  (List.length l2, l2)
(*@ (s, r) = count_even l
    ensures s = List.length r
    ensures forall x: int. List.mem x r -> x > 0 && mod x 2 = 0 
    ensures forall x: int. List.mem x r -> List.mem x l
    ensures forall x: int. not List.mem x l -> not List.mem x r *)
\end{gospell}

\subsection*{Generated CakeML}

\begin{cakeml}
fun count_even l =
  let val l1 = l in
  let val l11 = filter (fn x => (x > (0)) andalso ((mod x (2)) = (0))) l1 in
  let val l2 = rev l11 in (List.length l2, l2)
\end{cakeml}

\subsection*{Corrected CakeML}

\begin{cakeml}
fun count_even l =
  let val l1 = List.filter (fn x => (x > (0)) andalso ((Int.mod x (2)) = (0))) l in
  let val l2 = List.rev l1 in (List.length l2, l2) end end
\end{cakeml}

\newpage

\section*{Depth-search tree}

\subsection*{OCaml + GOSPEL}

\begin{gospell}
(*@ function rec to_list (t: 'a tree) : 'a list = 
  match t with
  | Leaf -> []
  | Node l x r -> x :: to_list l @ to_list r
*)
(*@
  variant t
*)

let rec depth_search t n = 
  match t with
  | Leaf -> false
  | Node (l,x,r) -> x = n || depth_search l n || depth_search r n
(*@
  r = depth_search t n
  variant t
  ensures List.mem n (to_list t) <-> r
*)
\end{gospell}

\subsection*{Generated CakeML}

\begin{cakeml}
'a datatype tree = Leaf | Node of 'a tree * 'a * 'a tree

fun depth_search t n = let val t1 = t in
  let val n1 = n in
  (case t1 of
    Leaf => false
  | Node l x r =>
    (x = n1) orelse ((depth_search l n1) orelse (depth_search r n1)))
\end{cakeml}

\subsection*{Corrected CakeML}

\begin{cakeml}
datatype 'a tree = Leaf | Node ('a tree) 'a ('a tree)

fun depth_search t n =
  (case t of
    Leaf => False
  | Node l x r =>
    (x = n) orelse ((depth_search l n) orelse (depth_search r n)))
\end{cakeml}

\newpage

\section*{Tree Comparison}

\subsection*{OCaml + GOSPEL}

\begin{gospell}
module Tree = struct
  type 'a tree =
    | Leaf
    | Node of 'a tree * 'a * 'a tree

  let rec cmp t1 t2 =
    match t1, t2 with
    | Leaf, Leaf -> true
    | Leaf, _ -> false
    | _, Leaf -> false
    | Node (l1,x1,r1), Node (l2,x2,r2) -> cmp l1 l2 && x1 = x2 && cmp r1 r2
  (*@
  r = cmp t1 t2
  variant t1
  ensures r <-> t1 = t2
  *)
end
\end{gospell}

\subsection*{Generated CakeML}

\begin{cakeml}
'a datatype tree = Leaf | Node of 'a tree * 'a * 'a tree

fun cmp t1 t2 = let val t11 = t1 in
  let val t21 = t2 in
  (case (t11, t21) of
    (Leaf, Leaf) => true
  | (Leaf, _) => false
  | (_, Leaf) => false
  | (Node l1 x1 r1, Node l2 x2 r2) =>
    (cmp l1 l2) andalso ((x1 = x2) andalso (cmp r1 r2)))
\end{cakeml}

\subsection*{Corrected CakeML}

\begin{cakeml}
structure Tree = struct
  datatype 'a tree = Leaf | Node ('a tree) 'a ('a tree)

  fun cmp t1 t2 =
  (case (t1, t2) of
    (Leaf, Leaf) => True
  | (Leaf, _) => False
  | (_, Leaf) => False
  | (Node l1 x1 r1, Node l2 x2 r2) =>
      (cmp l1 l2) andalso ((x1 = x2) andalso (cmp r1 r2)))
end
\end{cakeml}

\newpage

\section*{References}

\subsection*{OCaml + GOSPEL}

\begin{gospell}
let rec eudiv_aux (x: int) (y: int) (r: int ref) (q: int ref) =
  if !r >= y then (r := !r - y; q := !q + 1; eudiv_aux x y r q)
  else ()
(*@ res = eudiv_aux x y r q
    requires x >= 0
    requires y > 0
    requires 0 <= !r <= x
    requires x = y * !q + !r
    variant !r - y
    ensures x = y * !q + !r
    ensures 0 <= !r <= x
    ensures 0 <= !r < y *)

let euclidean_div x y =
    let r = ref x and q = ref 0 in 
  eudiv_aux x y r q;
  (!q, !r)
(*@ (q, r) = euclidean_div x y
    requires x >= 0
    requires y > 0
    ensures x = y * q + r
    ensures 0 <= r < y *)
\end{gospell}

\subsection*{Generated CakeML}

\begin{cakeml}
fun eudiv_aux x y r q = let val x1 = x in
  let val y1 = y in
  let val r1 = r in
  let val q1 = q in
  if (!r1) >= y1
  then ((r1 := ((!r1) - y1); q1 := ((!q1) + (1)); eudiv_aux x1 y1 r1 q1))

fun euclidean_div x y =
  let val x1 = x in
  let val y1 = y in
  let val r = ref x1 in
  let val q = ref (0) in (eudiv_aux x1 y1 r q; (!q, !r))
\end{cakeml}

\newpage

\subsection*{Corrected CakeML}

\begin{cakeml}
fun eudiv_aux x y r q =
  if (!r) >= y
  then ((r := ((!r) - y); q := ((!q) + (1)); eudiv_aux x y r q))
  else ()

fun euclidean_div x y =
  let val r = Ref x in
  let val q = Ref (0) in (eudiv_aux x y r q; (!q, !r)) 
    end end
\end{cakeml}

\section{Red Black Trees}

\subsection{Verified \whyml Code}
\label{subsec:Red_Black_Trees_Whyml_Og}

\begin{whylang}
module RedBlackTree

  (* Red-black trees (data type) *)

  type key = int
  type value = int

  type color = Red | Black

  type tree =
    | Leaf
    | Node color tree key value tree

  (* occurrence of a key/value pair in a tree *)

  predicate memt (t : tree) (k : key) (v : value) =
    match t with
    | Leaf -> false
    | Node _ l k' v' r -> (k = k' /\ v = v') \/ memt l k v \/ memt r k v
    end

  lemma memt_color :
    forall l r : tree, k k' : key, v v' : value, c c' : color.
    memt (Node c l k v r) k' v' -> memt (Node c' l k v r) k' v'

  (* binary search tree *)

  use int.Int

  predicate lt_tree (x : key) (t : tree) =
    forall k : key. forall v : value. memt t k v -> k < x

  predicate gt_tree (x : key) (t : tree) =
    forall k : key. forall v : value. memt t k v -> x < k

  lemma lt_leaf: forall x: key. lt_tree x Leaf

  lemma gt_leaf: forall x: key. gt_tree x Leaf

  lemma lt_tree_node:
    forall x y: key, v: value, l r: tree, c: color.
    lt_tree x l -> lt_tree x r -> y < x -> lt_tree x (Node c l y v r)

  lemma gt_tree_node:
    forall x y: key, v: value, l r: tree, c: color.
    gt_tree x l -> gt_tree x r -> x < y -> gt_tree x (Node c l y v r)

  lemma lt_node_lt:
    forall x y: key, v: value, l r: tree, c: color.
    lt_tree x (Node c l y v r) -> y < x

  lemma gt_node_gt:
    forall x y: key, v: value, l r: tree, c: color.
    gt_tree x (Node c l y v r) -> x < y

  lemma lt_left:
    forall x y: key, v: value, l r: tree, c : color.
    lt_tree x (Node c l y v r) -> lt_tree x l

  lemma lt_right:
    forall x y: key, v: value,  l r: tree, c: color.
    lt_tree x (Node c l y v r) -> lt_tree x r

  lemma gt_left:
    forall x y: key, v: value, l r: tree, c: color.
    gt_tree x (Node c l y v r) -> gt_tree x l

  lemma gt_right:
    forall x y: key, v: value, l r: tree, c: color.
    gt_tree x (Node c l y v r) -> gt_tree x r

  lemma lt_tree_not_in:
    forall x: key, t: tree. lt_tree x t -> forall v: value. not (memt t x v)

  lemma lt_tree_trans:
    forall x y: key. x < y -> forall t: tree. lt_tree x t -> lt_tree y t

  lemma gt_tree_not_in :
    forall x: key, t: tree. gt_tree x t -> forall v: value. not (memt t x v)

  lemma gt_tree_trans :
    forall x y: key. y < x -> forall t: tree. gt_tree x t -> gt_tree y t

  predicate bst (t : tree) =
    match t with
    | Leaf -> true
    | Node _ l k _ r -> bst l /\ bst r /\ lt_tree k l /\ gt_tree k r
    end

  lemma bst_Leaf : bst Leaf

  lemma bst_left:
    forall k: key, v: value, l r: tree, c: color. bst (Node c l k v r) -> bst l

  lemma bst_right:
    forall k: key, v: value, l r: tree, c: color. bst (Node c l k v r) -> bst r

  lemma bst_color:
    forall c c': color, k: key, v: value, l r: tree.
    bst (Node c l k v r) -> bst (Node c' l k v r)

  lemma rotate_left:
    forall kx ky: key, vx vy: value, a b c: tree, c1 c2 c3 c4: color.
    bst (Node c1 a kx vx (Node c2 b ky vy c)) ->
    bst (Node c3 (Node c4 a kx vx b) ky vy c)

  lemma rotate_right:
    forall kx ky: key, vx vy: value, a b c: tree, c1 c2 c3 c4: color.
    bst (Node c3 (Node c4 a kx vx b) ky vy c) ->
    bst (Node c1 a kx vx (Node c2 b ky vy c))

  predicate is_not_red (t : tree) =
    match t with
    | Node Red _ _ _ _ -> false
    | Leaf | Node Black _ _ _ _ -> true
    end

  predicate rbtree (n : int) (t : tree) =
    match t with
    | Leaf ->
        n = 0
    | Node Red   l _ _ r ->
        rbtree n l /\ rbtree n r /\ is_not_red l /\ is_not_red r
    | Node Black l _ _ r ->
        rbtree (n-1) l /\ rbtree (n-1) r
    end

  lemma rbtree_Leaf:
    rbtree 0 Leaf

  lemma rbtree_Node1:
    forall k:key, v:value. rbtree 0 (Node Red Leaf k v Leaf)

  lemma rbtree_left:
    forall x: key, v: value, l r: tree, c: color.
    (exists n: int. rbtree n (Node c l x v r)) -> exists n: int. rbtree n l

  lemma rbtree_right:
    forall x: key, v: value, l r: tree, c: color.
    (exists n: int. rbtree n (Node c l x v r)) -> exists n: int. rbtree n r

  exception Not_found

  let rec find (t : tree) (k : key) : value
    requires { bst t }
    variant { t }
    ensures { memt t k result }
    raises { Not_found -> forall v : value. not (memt t k v) }
  = match t with
    | Leaf -> raise Not_found
    | Node _ l k' v r ->
        if k = k' then v
        else if k < k' then find l k
        else (* k > k' *) find r k
    end

  predicate almost_rbtree (n : int) (t : tree) =
    match t with
    | Leaf ->
        n = 0
    | Node Red   l _ _ r ->
        rbtree n l /\ rbtree n r
    | Node Black l _ _ r ->
        rbtree (n-1) l /\ rbtree (n-1) r
    end

  lemma rbtree_almost_rbtree:
    forall n: int, t: tree. rbtree n t -> almost_rbtree n t

  lemma rbtree_almost_rbtree_ex:
    forall s: tree.
    (exists n: int. rbtree n s) -> exists n: int. almost_rbtree n s

  lemma almost_rbtree_rbtree_black:
    forall x: key, v: value, l r: tree, n: int.
    almost_rbtree n (Node Black l x v r) -> rbtree n (Node Black l x v r)

  let lbalance (l : tree) (k : key) (v : value) (r : tree)
    requires { lt_tree k l /\ gt_tree k r /\ bst l /\ bst r }
    ensures { bst result /\
      (forall n : int. almost_rbtree n l -> rbtree n r -> rbtree (n+1) result)
      /\
      forall k':key, v':value.
        memt result k' v' <->
        if k' = k then v' = v else (memt l k' v' \/ memt r k' v') }
  = match l with
    | Node Red (Node Red a kx vx b) ky vy c
    | Node Red a kx vx (Node Red b ky vy c) ->
        Node Red (Node Black a kx vx b) ky vy (Node Black c k v r)
    | _ ->
        Node Black l k v r
    end

  let rbalance (l : tree) (k : key) (v : value) (r : tree)
    requires { lt_tree k l /\ gt_tree k r /\ bst l /\ bst r }
    ensures { bst result /\
      (forall n : int. almost_rbtree n r -> rbtree n l -> rbtree (n+1) result)
      /\
      forall k':key, v':value.
        memt result k' v' <->
        if k' = k then v' = v else (memt l k' v' \/ memt r k' v') }
  = match r with
    | Node Red (Node Red b ky vy c) kz vz d
    | Node Red b ky vy (Node Red c kz vz d) ->
        Node Red (Node Black l k v b) ky vy (Node Black c kz vz d)
    | _ ->
        Node Black l k v r
    end

  (* `insert x s` inserts `x` in tree `s`, resulting in a possible top red-red
     conflict when `s` is red. *)

  let rec insert (t : tree) (k : key) (v : value) : tree
    requires { bst t /\ exists n: int. rbtree n t }
    variant { t }
    ensures { bst result /\
      (forall n : int. rbtree n t -> almost_rbtree n result /\
         (is_not_red t -> rbtree n result)) /\
      memt result k v /\
      forall k':key, v':value.
      memt result k' v' <-> if k' = k then v' = v else memt t k' v' }
  = match t with
    | Leaf ->
        Node Red Leaf k v Leaf
    | Node Red l k' v' r ->
        if k < k' then Node Red (insert l k v) k' v' r
        else if k' < k then Node Red l k' v' (insert r k v)
        else (* k = k' *) Node Red l k' v r
    | Node Black l k' v' r ->
        if k < k' then lbalance (insert l k v) k' v' r
        else if k' < k then rbalance l k' v' (insert r k v)
        else (* k = k' *) Node Black l k' v r
    end

  (* finally `add x s` calls `insert` and recolors the root as black *)

  let add (t : tree) (k : key) (v : value) : tree
    requires { bst t /\ exists n:int. rbtree n t }
    ensures { bst result /\ (exists n:int. rbtree n result) /\
      memt result k v /\
      forall k':key, v':value.
      memt result k' v' <-> if k' = k then v' = v else memt t k' v' }
  = match insert t k v with
    | Node _ l k' v' r -> Node Black l k' v' r
    | Leaf -> absurd
    end

end
\end{whylang}

\subsection{Generated \cml Code From Extraction}
\label{subsec:Red_Black_Trees_Cakeml}

\begin{cakeml}
type key = int

type value = int

datatype color = Red | Black

datatype tree = Leaf | Node color tree (int) (int) tree

exception Not_found



fun find t k =
  (case t of
    Leaf => raise Not_found
  | (Node _ l kqt v r) =>
    if k = kqt then (v)  else (if k < kqt then (find l k)  else (find r k)))

fun lbalance l k v r =
  (case l of
    (Node Red (Node Red a kx vx b) ky vy c) | (Node Red a kx vx (Node Red b
    ky vy c)) =>
    Node (Red) (Node (Black) a kx vx b) ky vy (Node (Black) c k v r)
  | _ => Node (Black) l k v r)

fun rbalance l k v r =
  (case r of
    (Node Red (Node Red b ky vy c) kz vz d) | (Node Red b ky vy (Node Red c
    kz vz d)) =>
    Node (Red) (Node (Black) l k v b) ky vy (Node (Black) c kz vz d)
  | _ => Node (Black) l k v r)

fun insert t k v =
  (case t of
    Leaf => Node (Red) (Leaf) k v (Leaf)
  | (Node Red l kqt vqt r) =>
    if k < kqt then (Node (Red) (insert l k v) kqt vqt r) 
    else (if kqt < k then (Node (Red) l kqt vqt (insert r k v)) 
          else (Node (Red) l kqt v r))
  | (Node Black l kqt vqt r) =>
    if k < kqt then (lbalance (insert l k v) kqt vqt r) 
    else (if kqt < k then (rbalance l kqt vqt (insert r k v)) 
          else (Node (Black) l kqt v r)))

exception Absurd

fun add t k v =
  (case insert t k v of
    (Node _ l kqt vqt r) => Node (Black) l kqt vqt r
  | Leaf => raise Absurd)
\end{cakeml}

\subsection{Generated \cml Code + \gospel Annotations}
\label{subsec:Red_Black_Trees_Cakeml_Gospel}

\begin{cakeml}
type key = int
type value = int
datatype color = Red | Black
datatype tree = Leaf | Node color tree (int) (int) tree

(*@ predicate memt (t: tree) (k: key) (v: value) =
    match t with
    | Leaf -> false
    | Node _ l k' v' r -> (k = k' && v = v') || memt l k v || memt r k v *)

(*@ lemma memt_color :
    forall l r : tree, k k' : key, v v' : value, c c' : color.
    memt (Node c l k v r) k' v' -> memt (Node c' l k v r) k' v' *)

(*@ predicate lt_tree (x : key) (t : tree) =
    forall k : key. forall v : value. memt t k v -> k < x *)

(*@ predicate gt_tree (x : key) (t : tree) =
    forall k : key. forall v : value. memt t k v -> x < k *)

(*@ lemma lt_leaf: forall x: key. lt_tree x Leaf *)

(*@ lemma gt_leaf: forall x: key. gt_tree x Leaf *)

(*@ lemma lt_tree_node:
    forall x y: key, v: value, l r: tree, c: color.
      lt_tree x l -> lt_tree x r -> y < x -> lt_tree x (Node c l y v r) *)

(*@ lemma gt_tree_node:
    forall x y: key, v: value, l r: tree, c: color.
      gt_tree x l -> gt_tree x r -> x < y -> gt_tree x (Node c l y v r) *)

(*@ lemma lt_node_lt:
    forall x y: key, v: value, l r: tree, c: color.
      lt_tree x (Node c l y v r) -> y < x *)

(*@ lemma gt_node_gt:
    forall x y: key, v: value, l r: tree, c: color.
      gt_tree x (Node c l y v r) -> x < y *)

(*@ lemma lt_left:
    forall x y: key, v: value, l r: tree, c: color.
      lt_tree x (Node c l y v r) -> lt_tree x l *)

(*@ lemma lt_right:
    forall x y: key, v: value, l r: tree, c: color.
      lt_tree x (Node c l y v r) -> lt_tree x r *)

(*@ lemma gt_left:
    forall x y: key, v: value, l r: tree, c: color.
      gt_tree x (Node c l y v r) -> gt_tree x l *)

(*@ lemma gt_right:
    forall x y: key, v: value, l r: tree, c: color.
      gt_tree x (Node c l y v r) -> gt_tree x r *)

(*@ lemma lt_tree_not_in:
    forall x: key, t: tree. lt_tree x t -> forall v: value. not (memt t x v) *)

(*@ lemma lt_tree_trans:
    forall x y: key. x < y -> forall t: tree. lt_tree x t -> lt_tree y t *)

(*@ lemma gt_tree_not_in:
    forall x: key, t: tree. gt_tree x t -> forall v: value. not (memt t x v) *)

(*@ lemma gt_tree_trans:
    forall x y: key. y < x -> forall t: tree. gt_tree x t -> gt_tree y t *)

(*@ predicate bst (t: tree) =
    match t with
    | Leaf -> true
    | Node _ l k _ r -> bst l && bst r && lt_tree k l && gt_tree k r *)

(*@ lemma bst_Leaf: bst Leaf *)

(*@ lemma bst_left:
    forall k: key, v: value, l r: tree, c: color.
      bst (Node c l k v r) -> bst l *)

(*@ lemma bst_right:
    forall k: key, v: value, l r: tree, c: color.
      bst (Node c l k v r) -> bst r *)

(*@ lemma bst_color:
    forall c c': color, k: key, v: value, l r: tree.
      bst (Node c l k v r) -> bst (Node c' l k v r) *)

(*@ lemma rotate_left:
    forall kx ky: key, vx vy: value, a b c: tree, c1 c2 c3 c4: color.
      bst (Node c1 a kx vx (Node c2 b ky vy c)) ->
      bst (Node c3 (Node c4 a kx vx b) ky vy c) *)

(*@ lemma rotate_right:
    forall kx ky: key, vx vy: value, a b c: tree, c1 c2 c3 c4: color.
      bst (Node c3 (Node c4 a kx vx b) ky vy c) ->
      bst (Node c1 a kx vx (Node c2 b ky vy c)) *)

(*@ predicate is_not_red (t: tree) =
    match t with
    | Node Red _ _ _ _ -> false
    | Leaf -> true
    | Node Black _ _ _ _ -> true *)

(*@ predicate rbtree (n: int) (t: tree) =
    match t with
    | Leaf -> n = 0
    | Node Red l _ _ r ->
        rbtree n l && rbtree n r && is_not_red l && is_not_red r
    | Node Black l _ _ r ->
        rbtree (n-1) l && rbtree (n-1) r *)

(*@ lemma rbtree_Leaf: rbtree 0 Leaf *)

(*@ lemma rbtree_Node1:
    forall k: key, v: value. rbtree 0 (Node Red Leaf k v Leaf) *)

(*@ lemma rbtree_left:
    forall x: key, v: value, l r: tree, c: color.
      (exists n: int. rbtree n (Node c l x v r)) -> exists n: int. rbtree n l *)

(*@ lemma rbtree_right:
    forall x: key, v: value, l r: tree, c: color.
      (exists n: int. rbtree n (Node c l x v r)) -> exists n: int. rbtree n r *)

exception Not_found

fun find t k =
  (case t of
    Leaf => raise Not_found
  | (Node _ l kqt v r) =>
    if k = kqt then v
    else if k < kqt then find l k
    else find r k)
(*@ result = find t k
    requires bst t
    variant t
    ensures memt t k result
    raises Not_found -> forall v: value. not (memt t k v) *)

(*@ predicate almost_rbtree (n: int) (t: tree) =
    match t with
    | Leaf -> n = 0
    | Node Red l _ _ r -> rbtree n l /\ rbtree n r
    | Node Black l _ _ r -> rbtree (n-1) l /\ rbtree (n-1) r *)

(*@ lemma rbtree_almost_rbtree:
    forall n: int, t: tree. rbtree n t -> almost_rbtree n t *)

(*@ lemma rbtree_almost_rbtree_ex:
    forall s: tree.
      (exists n: int. rbtree n s) -> exists n: int. almost_rbtree n s *)

(*@ lemma almost_rbtree_rbtree_black:
    forall x: key, v: value, l r: tree, n: int.
      almost_rbtree n (Node Black l x v r) -> rbtree n (Node Black l x v r) *)

fun lbalance l k v r =
  (case l of
    (Node Red (Node Red a kx vx b) ky vy c) =>
      Node Red (Node Black a kx vx b) ky vy (Node Black c k v r)
  | (Node Red a kx vx (Node Red b ky vy c)) =>
      Node Red (Node Black a kx vx b) ky vy (Node Black c k v r)
  | _ => Node Black l k v r)
(*@ result = lbalance l k v r
    requires lt_tree k l && gt_tree k r && bst l && bst r
    ensures bst result &&
      (forall n: int. almost_rbtree n l -> rbtree n r -> rbtree (n+1) result) &&
      forall k': key, v': value.
        memt result k' v' <->
        if k' = k then v' = v else memt l k' v' || memt r k' v' *)

fun rbalance l k v r =
  (case r of
    (Node Red (Node Red b ky vy c) kz vz d) =>
      Node Red (Node Black l k v b) ky vy (Node Black c kz vz d)
  | (Node Red b ky vy (Node Red c kz vz d)) =>
      Node Red (Node Black l k v b) ky vy (Node Black c kz vz d)
  | _ => Node Black l k v r)
(*@ result = rbalance l k v r
    requires lt_tree k l && gt_tree k r && bst l && bst r
    ensures bst result &&
      (forall n: int. almost_rbtree n r -> rbtree n l -> rbtree (n+1) result) &&
      forall k': key, v': value.
        memt result k' v' <->
        if k' = k then v' = v else memt l k' v' || memt r k' v' *)

fun insert t k v =
  (case t of
    Leaf => Node Red Leaf k v Leaf
  | (Node Red l kqt vqt r) =>
    if k < kqt then Node Red (insert l k v) kqt vqt r
    else if kqt < k then Node Red l kqt vqt (insert r k v)
    else Node Red l kqt v r
  | (Node Black l kqt vqt r) =>
    if k < kqt then lbalance (insert l k v) kqt vqt r
    else if kqt < k then rbalance l kqt vqt (insert r k v)
    else Node Black l kqt v r)
(*@ result = insert t k v
    requires bst t && exists n: int. rbtree n t
    variant t
    ensures bst result &&
      (forall n: int. rbtree n t ->
        almost_rbtree n result &&
        (is_not_red t -> rbtree n result)) &&
      memt result k v &&
      forall k': key, v': value.
        memt result k' v' <->
        if k' = k then v' = v else memt t k' v' *)

exception Absurd

fun add t k v =
  (case insert t k v of
    (Node _ l kqt vqt r) => Node Black l kqt vqt r
  | Leaf => raise Absurd)
(*@ result = add t k v
    requires bst t
    requires exists n: int. rbtree n t
    ensures bst result
    ensures exists n: int. rbtree n result
    ensures memt result k v
    ensures forall k': key, v': value.
    memt result k' v' <-> if k' = k then v' = v else memt t k' v' *)
\end{cakeml}

\subsection{Generated \whyml Code}
\label{subsec:Red_Black_Trees_Whyml}

\begin{whylang}
module RedBlackTree

  use int.Int

  type key = int

  type value = int

  type color =
 | Red
 | Black

  type tree =
 | Leaf
 | Node color tree int int tree

  predicate memt (t: tree) (k: key) (v: value) =
 match t with
 | Leaf -> false
 | Node _ l k' v' r -> ((k = k') /\ (v = v')) \/ ((memt l k v) \/ (memt r k v))
  end

  lemma memt_color : forall l: tree, r: tree, k: key, k': key, v: value, v': value, c: color, c': color. (memt (Node c l k v r) k' v') -> (memt (Node c' l k v r) k' v')

  predicate lt_tree (x: key) (t: tree) =
 forall k: key. forall v: value. (memt t k v) -> (k < x)

  predicate gt_tree (x: key) (t: tree) =
 forall k: key. forall v: value. (memt t k v) -> (x < k)

  lemma lt_leaf : forall x: key. lt_tree x Leaf

  lemma gt_leaf : forall x: key. gt_tree x Leaf

  lemma lt_tree_node : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (lt_tree x l) -> ((lt_tree x r) -> ((y < x) -> (lt_tree x (Node c l y v r))))

  lemma gt_tree_node : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (gt_tree x l) -> ((gt_tree x r) -> ((x < y) -> (gt_tree x (Node c l y v r))))

  lemma lt_node_lt : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (lt_tree x (Node c l y v r)) -> (y < x)

  lemma gt_node_gt : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (gt_tree x (Node c l y v r)) -> (x < y)

  lemma lt_left : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (lt_tree x (Node c l y v r)) -> (lt_tree x l)

  lemma lt_right : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (lt_tree x (Node c l y v r)) -> (lt_tree x r)

  lemma gt_left : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (gt_tree x (Node c l y v r)) -> (gt_tree x l)

  lemma gt_right : forall x: key, y: key, v: value, l: tree, r: tree, c: color. (gt_tree x (Node c l y v r)) -> (gt_tree x r)

  lemma lt_tree_not_in : forall x: key, t: tree. (lt_tree x t) -> (forall v: value. not memt t x v)

  lemma lt_tree_trans : forall x: key, y: key. (x < y) -> (forall t: tree. (lt_tree x t) -> (lt_tree y t))

  lemma gt_tree_not_in : forall x: key, t: tree. (gt_tree x t) -> (forall v: value. not memt t x v)

  lemma gt_tree_trans : forall x: key, y: key. (y < x) -> (forall t: tree. (gt_tree x t) -> (gt_tree y t))

  predicate bst (t: tree) =
 match t with
 | Leaf -> true
 | Node _ l k _ r -> (bst l) /\ ((bst r) /\ ((lt_tree k l) /\ (gt_tree k r)))
  end

  lemma bst_Leaf : bst Leaf

  lemma bst_left : forall k: key, v: value, l: tree, r: tree, c: color. (bst (Node c l k v r)) -> (bst l)

  lemma bst_right : forall k: key, v: value, l: tree, r: tree, c: color. (bst (Node c l k v r)) -> (bst r)

  lemma bst_color : forall c: color, c': color, k: key, v: value, l: tree, r: tree. (bst (Node c l k v r)) -> (bst (Node c' l k v r))

  lemma rotate_left : forall kx: key, ky: key, vx: value, vy: value, a: tree, b: tree, c: tree, c1: color, c2: color, c3: color, c4: color. (bst (Node c1 a kx vx (Node c2 b ky vy c))) -> (bst (Node c3 (Node c4 a kx vx b) ky vy c))

  lemma rotate_right : forall kx: key, ky: key, vx: value, vy: value, a: tree, b: tree, c: tree, c1: color, c2: color, c3: color, c4: color. (bst (Node c3 (Node c4 a kx vx b) ky vy c)) -> (bst (Node c1 a kx vx (Node c2 b ky vy c)))

  predicate is_not_red (t: tree) =
 match t with
 | Node Red _ _ _ _ -> false
 | Leaf -> true
 | Node Black _ _ _ _ -> true
  end

  predicate rbtree (n: int) (t: tree) =
 match t with
 | Leaf -> n = 0
 | Node Red l _ _ r -> (rbtree n l) /\ ((rbtree n r) /\ ((is_not_red l) /\ (is_not_red r)))
 | Node Black l _ _ r -> (rbtree (n - 1) l) /\ (rbtree (n - 1) r)
  end

  lemma rbtree_Leaf : rbtree 0 Leaf

  lemma rbtree_Node1 : forall k: key, v: value. rbtree 0 (Node Red Leaf k v Leaf)

  lemma rbtree_left : forall x: key, v: value, l: tree, r: tree, c: color. (exists n: int. rbtree n (Node c l x v r)) -> (exists n: int. rbtree n l)

  lemma rbtree_right : forall x: key, v: value, l: tree, r: tree, c: color. (exists n: int. rbtree n (Node c l x v r)) -> (exists n: int. rbtree n r)

  exception Not_found

  let rec find t k
  requires { bst t }
  variant { t }
  ensures { memt t k result }
  raises { Not_found () -> forall v: value. not memt t k v } 
 = match t with
  | Leaf -> raise Not_found
  | Node _ l kqt v r -> if k = kqt then v else if k < kqt then find l k else find r k
  end

  predicate almost_rbtree (n: int) (t: tree) =
 match t with
 | Leaf -> n = 0
 | Node Red l _ _ r -> (rbtree n l) /\ (rbtree n r)
 | Node Black l _ _ r -> (rbtree (n - 1) l) /\ (rbtree (n - 1) r)
  end

  lemma rbtree_almost_rbtree : forall n: int, t: tree. (rbtree n t) -> (almost_rbtree n t)

  lemma rbtree_almost_rbtree_ex : forall s: tree. (exists n: int. rbtree n s) -> (exists n: int. almost_rbtree n s)

  lemma almost_rbtree_rbtree_black : forall x: key, v: value, l: tree, r: tree, n: int. (almost_rbtree n (Node Black l x v r)) -> (rbtree n (Node Black l x v r))

  let lbalance l k v r
  requires { (lt_tree k l) /\ ((gt_tree k r) /\ ((bst l) /\ (bst r))) }
  ensures { (bst result) /\ ((forall n: int. (almost_rbtree n l) -> ((rbtree n r) -> (rbtree (n + 1) result))) /\ (forall k': key, v': value. (memt result k' v') <-> (if k' = k then v' = v else (memt l k' v') \/ (memt r k' v')))) } 
 = match l with
  | Node Red (Node Red a kx vx b) ky vy c -> Node Red (Node Black a kx vx b) ky vy (Node Black c k v r)
  | Node Red a kx vx (Node Red b ky vy c) -> Node Red (Node Black a kx vx b) ky vy (Node Black c k v r)
  | _ -> Node Black l k v r
  end

  let rbalance l k v r
  requires { (lt_tree k l) /\ ((gt_tree k r) /\ ((bst l) /\ (bst r))) }
  ensures { (bst result) /\ ((forall n: int. (almost_rbtree n r) -> ((rbtree n l) -> (rbtree (n + 1) result))) /\ (forall k': key, v': value. (memt result k' v') <-> (if k' = k then v' = v else (memt l k' v') \/ (memt r k' v')))) } 
 = match r with
  | Node Red (Node Red b ky vy c) kz vz d -> Node Red (Node Black l k v b) ky vy (Node Black c kz vz d)
  | Node Red b ky vy (Node Red c kz vz d) -> Node Red (Node Black l k v b) ky vy (Node Black c kz vz d)
  | _ -> Node Black l k v r
  end

  let rec insert t k v
  requires { (bst t) /\ (exists n: int. rbtree n t) }
  variant { t }
  ensures { (bst result) /\ ((forall n: int. (rbtree n t) -> ((almost_rbtree n result) /\ ((is_not_red t) -> (rbtree n result)))) /\ ((memt result k v) /\ (forall k': key, v': value. (memt result k' v') <-> (if k' = k then v' = v else memt t k' v')))) } 
 = match t with
  | Leaf -> Node Red Leaf k v Leaf
  | Node Red l kqt vqt r -> if k < kqt then Node Red (insert l k v) kqt vqt r else if kqt < k then Node Red l kqt vqt (insert r k v) else Node Red l kqt v r
  | Node Black l kqt vqt r -> if k < kqt then lbalance (insert l k v) kqt vqt r else if kqt < k then rbalance l kqt vqt (insert r k v) else Node Black l kqt v r
  end

  

  let add t k v
  requires { bst t }
  requires { exists n: int. rbtree n t }
  ensures { bst result }
  ensures { exists n: int. rbtree n result }
  ensures { memt result k v }
  ensures { forall k': key, v': value. (memt result k' v') <-> (if k' = k then v' = v else memt t k' v') } 
 = match insert t k v with
  | Node _ l kqt vqt r -> Node Black l kqt vqt r
  | Leaf -> absurd
  end

end
\end{whylang}

\newcommand{\explanation}[1]{\cellcolor{green!20}#1}
\newcommand{\noresult}{\cellcolor{gray!20}---}
\newcommand{\blank}{\cellcolor{gray!20}}
\newcommand{\valid}[1]{\cellcolor{green!20}#1}

\section{Prover Performance Tables}
\label{sec:Prover_Performance_Tables}

\subsection{Red Black Trees Original \whyml}
\label{subsec:Red_Black_Trees_Original_Whyml}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\ 
\hline
\explanation{memt\_color} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{lt\_leaf} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{gt\_leaf} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{lt\_tree\_node} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{gt\_tree\_node} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{lt\_node\_lt} & \noresult& \noresult& \noresult& \valid{0.03} & \noresult \\ 
\hline
\explanation{gt\_node\_gt} & \noresult& \noresult& \noresult& \valid{0.03} & \noresult \\ 
\hline
\explanation{lt\_left} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{lt\_right} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{gt\_left} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{gt\_right} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{lt\_tree\_not\_in} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{lt\_tree\_trans} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{gt\_tree\_not\_in} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{gt\_tree\_trans} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{bst\_Leaf} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{bst\_left} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{bst\_right} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{bst\_color} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{rotate\_left} & \valid{0.02} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{rotate\_right} & \valid{0.02} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{rbtree\_Leaf} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{rbtree\_Node1} & \noresult& \noresult& \noresult& \noresult& \valid{0.01} \\ 
\hline
\explanation{rbtree\_left} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{rbtree\_right} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{VC for find} & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\explanation{rbtree\_almost\_rbtree} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{rbtree\_almost\_rbtree\_ex} & \valid{0.01} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{almost\_rbtree\_rbtree\_black} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{VC for lbalance} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{0.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\end{tabular}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\  
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.06} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\explanation{VC for rbalance}  & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{0.5em}\explanation{postcondition}& \noresult& \noresult&  \noresult&  \noresult& \noresult \\ 
\hline
\hspace{0.5em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1em}\explanation{postcondition}& \noresult& \noresult&  \noresult&  \noresult& \noresult \\ 
\hline
\end{tabular}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\
\hline
\hspace{1em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.05} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.04} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \valid{0.80} & \noresult& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.04} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\explanation{VC for insert} & \valid{0.50} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{VC for add} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{0.5em}\explanation{precondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{unreachable point} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{destruct\_term o} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{introduce\_exists} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{destruct\_term x4} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{2em}\explanation{postcondition} & \noresult& \noresult& \valid{0.03}& \noresult& \noresult\\ 
\hline
\hspace{2em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{2em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{2.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{exists(n+1)} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{3em}\explanation{add'vc.2.0.0.1.1.0} &\noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{postcondition} &\noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\end{tabular}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\
\hline
\hspace{2.5em}\explanation{postcondition} &\noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{postcondition} &\noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{postcondition} &\noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline \end{tabular}

\subsection{Red Black Trees Translated \whyml}
\label{subsec:Red_Black_Trees_Translated_Whyml}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\ 
\hline
\explanation{memt\_color} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{lt\_leaf} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{gt\_leaf} & \noresult& \noresult& \valid{0.00}& \noresult& \noresult \\ 
\hline
\explanation{lt\_tree\_node} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{gt\_tree\_node} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{lt\_node\_lt} & \noresult& \noresult& \noresult& \valid{0.02} & \noresult \\ 
\hline
\explanation{gt\_node\_gt} & \noresult& \noresult& \noresult& \valid{0.03} & \noresult \\ 
\hline
\explanation{lt\_left} & \noresult& \noresult& \noresult& \noresult& \valid{0.00} \\ 
\hline
\explanation{lt\_right} & \noresult& \noresult& \noresult& \noresult& \valid{0.00} \\ 
\hline
\explanation{gt\_left} & \noresult& \noresult& \noresult& \noresult& \valid{0.00} \\ 
\hline
\explanation{gt\_right} & \noresult& \noresult& \noresult& \noresult& \valid{0.00} \\ 
\hline
\explanation{lt\_tree\_not\_in} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{lt\_tree\_trans} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{gt\_tree\_not\_in} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{gt\_tree\_trans} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{bst\_Leaf} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{bst\_left} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{bst\_right} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{bst\_color} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{rotate\_left} & \valid{0.02} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{rotate\_right} & \valid{0.02} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{rbtree\_Leaf} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{rbtree\_Node1} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult \\ 
\hline
\explanation{rbtree\_left} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{rbtree\_right} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{VC for find} & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\explanation{rbtree\_almost\_rbtree} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{rbtree\_almost\_rbtree\_ex} & \valid{0.01} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{almost\_rbtree\_rbtree\_black} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\explanation{VC for lbalance} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{0.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\end{tabular}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\  
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.04} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.10} & \noresult& \noresult\\
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.04} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.03} & \noresult& \noresult\\ 
\hline
\explanation{VC for rbalance}  & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{0.5em}\explanation{postcondition}& \noresult& \noresult&  \noresult&  \noresult& \noresult \\ 
\hline
\hspace{0.5em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1em}\explanation{postcondition}& \noresult& \noresult&  \noresult&  \noresult& \noresult \\ 
\hline
\hspace{1em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition}  & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.04} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \valid{0.03} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\end{tabular}

\begin{tabular}{|l|c|c|c|c|c|}
\hline Proof obligations & Alt-Ergo 2.6.2 & CVC4 1.8 & CVC5 1.2.1 & Spass 3.9 & Z3 4.15.3 \\
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.05} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult & \valid{0.69}& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.05} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.04} & \noresult& \noresult\\ 
\hline
\explanation{VC for insert} & \valid{0.48} & \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{VC for add} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{0.5em}\explanation{precondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{unreachable point} & \noresult& \noresult& \valid{0.02} & \noresult& \noresult\\
\hline
\hspace{0.5em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01}& \noresult& \noresult\\  
\hline
\hspace{0.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{destruct\_term o} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{introduce\_exists} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{1.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{1.5em}\explanation{destruct\_term x4} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{2em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{2em}\explanation{split\_vc} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{2.5em}\explanation{postcondition} & \noresult& \noresult& \noresult& \noresult& \noresult\\ 
\hline
\hspace{2.5em}\explanation{exists(n+1)} & \blank& \blank& \blank& \blank& \blank\\ 
\hline
\hspace{3em}\explanation{add'vc.2.0.0.1.1.0} &\noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline
\hspace{1em}\explanation{postcondition} & \noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{postcondition} &\noresult& \noresult& \valid{0.01} & \noresult& \noresult\\ 
\hline
\hspace{0.5em}\explanation{postcondition} &\noresult& \noresult& \valid{0.02} & \noresult& \noresult\\ 
\hline \end{tabular}