%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter5.tex
%% NOVA thesis document file
%%
%% Chapter with lots of dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter5.tex}%

\chapter{Extraction}
\label{cha:Extraction}

In this chapter, we identify the inaccuracies and limitations found on \whythree's extraction mechanism for \cml at the beginning of
this study, and establish a subset of \whyml that is to be supported by the end of the study. This subset is based on common features
between the languages.

\section{Overview of Extraction-based verification}

The Extraction process, in the context of deductive program verification, refers to the process of deriving executable code from a verified implementation, in accordance
to its specification, in another language. It systematically eliminates all proof-oriented content that carries no computational meaning. Rather than re-verifying
correctness in the target language, extraction relies on the guarantees already established during the verification phase, delegating to the extraction mechanism the
responsibility of faithfully preserving the program's computational behaviour across the language boundary. 

In \whythree, this process is formalized through a dedicated ghost code erasure phase followed by a (target) language-specific driver, which together ensure that the
logical constructs of a \whyml program spanning both program-level annotations such as pre-conditions, post-conditions, invariants, and ghost variables, and theory-level
constructs such as predicates, lemmas, and axioms that support the proof obligations are cleanly removed before the remaining computational fragment is translated to the
target language. It is precisely this translation step that introduces the constraints examined in this chapter, since the target language must be capable of representing
the extracted fragment, the differences between \whyml and \cml impose a well-defined boundary on the subset of \whyml constructs that can be viably used within the pipeline.

\section{\whyml Subset}
\label{sec:Whyml_Subset}

When dealing with extraction or translation-based pipelines, we must take into consideration the differences and supported features between the source language and the target language. 
As such, we have established a subset of \whyml that is to be supported when extracting to \cml. This section discusses the limitations and supported features of \cml, in comparison 
to those found in \whyml, which we had to take into consideration when developing that subset.

\subsection{Limitations}

The integration of \cml into the verification pipeline introduces a set of constraints that must be carefully accounted for. Unlike \ocaml and \whyml, 
which belong to the ML language family but are designed with expressiveness and specification in mind, \cml was developed primarily with compiler correctness 
as its central concern, making its language design reflect the requirements of a formally verified compilation chain rather than general-purpose functional programming.
As a consequence, \cml supports a more restricted set of language constructs than either \ocaml or \whyml. This motivates the adoption 
of a specific subset of \whyml in the present work, one chosen to ensure that the programs expressed in the pipeline can be faithfully extracted to \cml 
without introducing constructs that fall outside its supported fragment, thereby preserving coherence across the full pipeline.

The subset was defined based on the limitations encountered during the extraction of \whyml code into \cml. The first restriction concerns the character 
\inlinecode{o}, which in \cml is strictly reserved for function composition and therefore cannot be used as an identifier for any variable or function name consisting 
solely of that character must be avoided within the subset. A further limitation arises with or-patterns since \cml does not support the grouping of multiple patterns under a 
single branch, meaning that constructs of the form \inlinecode{| E | N l x r -> ...} are not accepted and must instead be decomposed into separate branches, inevitably 
introducing code duplication. Additionally, \cml does not support imperative loop constructs such as \inlinecode{while} or \inlinecode{for}, requiring that all iterative 
behaviour be expressed through explicit recursion. Functors are likewise unsupported, and mutually recursive type definitions are not accepted by \cml, both of which are 
consequently excluded from the subset.

It should be noted that the subset defined in this chapter is not automatically enforced by the pipeline. No static analysis or syntactic check is currently performed to verify 
that a given \whyml program operates in line with the subset defined before extraction. The user is therefore expected to respect this subset throughout the pipeline in order 
to ensure correctness across both the extraction and translation phases of the study.

\subsection{Supported Features}

Despite the imposed restrictions, the resulting subset retains comprehensive support for variables and naming except \inlinecode{o},
an extended pattern matching for single branch patterns, wildcard patterns, constructor patterns, tuple patterns, list
patterns and nested patterns. In terms of the function there is still all the \whyml features such as regular function definitions,
recursive functions, anonymous functions and high-order functions. Data types continue to support algebraic data type definitions,
type aliases, polymorphic types and non-multiple recursive type definitions. Data structures like lists, tuples and references are supported,
like exception handling. Lastly, all the specifications of logical predicates, lemmas, axioms, ghost variables, invariants, pre-conditions and
post-conditions.

\section{Corrections to the \cml Extraction Driver}
\label{sec:Corrections_Cakeml}

The \cml extraction driver was already partially implemented within the Why3 platform, however it contained a number of inaccuracies that produced syntactically incorrect 
\cml output. The majory of corrections applied to the \cml extraction driver are syntactic in nature, addressing grammatical discrepancies between \whyml and \cml without
affecting the verification guarantees established during the proof phase. These include: capitalization conventions, let binding delimiters, and exception handling. Such
corrects were made to conform to the syntactic requirements of the \cml compiler. Computational semantics of the extracted program remain entirely unchanged. As described
in Section~\ref{sec:CakeML_Extraction}, extraction to \cml is invoked via \cameleer using the \inlinecode{--extract} flag. The following subsections detail each individual
change applied to the driver:

\subsection{Capitalization Conventions}

\cml requires that certain types and literals be capitalized as part of its syntactic grammar. The previous version of the extraction driver did 
not respect this convention, producing lowercase identifiers where uppercase is grammatically required. Notable examples include the boolean literals \inlinecode{true} 
and \inlinecode{false}, which must be rendered as \inlinecode{True} and \inlinecode{False} respectively, as well as reference construction, where \inlinecode{ref} must 
appear as \inlinecode{Ref}.

\subsection{Let Bindings Delimiters}

A syntactic distinction between \whyml and \cml concerns the delimitation of let bindings. In \whyml, let bindings do not require an explicit termination token, as the 
scope of the bound variable is determined by the structure of the expression. \cml, however, requires the \inlinecode{end} token to explicitly signal the closing of 
the binding scope. 

\begin{cakeml}
    let val a = 1
\end{cakeml} 

The correction applied to the extraction driver ensures that this token is consistently emitted at the appropriate position in the generated code.

\begin{cakeml}
    let val a = 1 end
\end{cakeml} 

\subsection{Data Type Declarations}

The declaration generated from the extraction was incorrect in the ordering and how the blocks were delimited. \cml does not split the blocks for the data types
with \inlinecode{*} as it only needs them to be correctly placed in blocks so it can be correctly written and also the token of is not necessary. The ordering only 
applies for the alpha where it should be in front of the data type instead of behind. As shown in the following example all these issues are present and need to be
fixed for correct \cml grammar.

\begin{cakeml}
    'a datatype tree = Leaf | Node of 'a tree * 'a * 'a tree
\end{cakeml}

The previous code shows clearly what needs to be changed in the printer and after applying the necessary modifications the code ends up as this.

\begin{cakeml}
    datatype 'a tree = Leaf | Node ('a tree) 'a ('a tree)
\end{cakeml}


\subsection{Exception Handling}

Exception handling introduces one of the more significant structural differences between \whyml and \cml. While \whyml permits exception declarations within local scopes 
using let bindings, \cml restricts exception declarations exclusively to the top level. To account for this, the extraction driver performs an initial pass over the function 
body to identify any exception declarations or raises, after which the declarations are hoisted to the top level and omitted from their original position in the generated code. 
Furthermore, the exceptions encountered during this work were not found in \cml's standard library, which appears to offer limited support for the exceptions commonly assumed 
by \whyml programs, requiring them to be explicitly declared in the generated output.

\begin{cakeml}
    fun search_aux a c n = 
    let exception Break of (int) in
    ((if c = (a.length) then (raise Not_found) 
        else (if (get a c) = n then (raise (Break n)) 
            else (search_aux a (c + (1)) n)))
    handle   Break i => i)
\end{cakeml}

The original code makes the exception declaration inside the scope of the function and that needs to be removed and placed outside of the scope.

\begin{cakeml}
    exception Not_found
    exception Break int

    fun search_aux a c n = 
    ((if c = (a.length) then (raise Not_found) 
        else (if (get a c) = n then (raise (Break n)) 
            else (search_aux a (c + (1)) n)))
    handle   Break i => i)
\end{cakeml}

A separate but related concern is the handling of the \inlinecode{absurd} construct. Since \cml provides no equivalent, a pragmatic decision was made to map \inlinecode{absurd} 
to an exception, generating an \inlinecode{exception Absurd} declaration at the top level and a corresponding \inlinecode{raise Absurd} at each point of occurrence. This 
represents the simplest viable solution for preserving the intent of the construct when it comes to marking a branch as unreachable within \cml's constraints.

The one notable exception is the handling of the \inlinecode{absurd} construct. In \whyml, \inlinecode{absurd} carries a precise logical meaning, it asserts that a given branch 
is unreachable, a claim that must be formally justified by a discharged proof obligation. Mapping this construct to \inlinecode{raise Absurd} in \cml is ultimately not a purely 
syntactic transformation, it replaces a statically verified impossibility with a runtime exception. While raising an exception is not a perfect semantic equivalent of an 
unreachable branch, it represents a practical solution within \cml's constraints, given that a correctly verified program should never reach that point at runtime.

\begin{cakeml}
    fun add t k v =
        (case insert t k v of
            (Node _ l kqt vqt r) => Node (Black) l kqt vqt r
        | Leaf => absurd)
\end{cakeml}

As much as the original code is correct for the way \inlinecode{absurd} is used, it is not acceptable in \cml so we opted for this solution.

\begin{cakeml}
    exception Absurd
    fun add t k v =
        (case insert t k v of
            (Node _ l kqt vqt r) => Node (Black) l kqt vqt r
        | Leaf => raise Absurd)
\end{cakeml}

While raising a runtime exception is not a direct representation of termination, it provides a simpler way to create the same effect.