%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter8.tex
%% NOVA thesis document file
%%
%% Chapter with lots of dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter8.tex}%

\chapter{Conclusion}
\label{cha:Conclusion}

On this work we focused on updating the extraction pipeline found in \whyml for \cml and creating a new translation tool in the opposite
direction while maintaining the program correctness and certified compilers workflow. In this chapter we present some final thoughts on this study,
the contributions for new tools in the improved pipeline and a few directions for future work.

\section{Final Thoughts}

For the final conclusions of this study it is important to answer the questions made in \ref{sec:Problem_Definition}. The results showed there is no clear
distinction in program correctness regarding the starting point of each approach, both reach the necessary state of verification when presented with
the necessary provers. The tools created ensured the correct adaptation of the \gospel specification language for the ML language family, showing that
this specification language can be integrated quite effortlessly without changing the main grammar of the ML language while still ensuring the necessary
specifications for discharging the invariants. Ultimately adding the translation tool changed the workflow of certified compilers in a way that there is now
a way to transform the language those compilers use in order to verify their content and ensuring not only the compiler process is safe from compiler
issues but also from the main code specifications.

In the end, this study gave an already existing library of \cml code a way to verify it with the addition of \gospel specifications.

\section{Contributions}

The first contributions of this work can be summarized as follows. The first contribution concerns the correction of the \cml extraction driver within the \whythree platform, which contained a 
number of inaccuracies that prevented the generation of syntactically correct \cml code. The corrections resulted in a driver capable of producing compilable \cml output from verified \whyml programs 
without manual intervention.

The second contribution is the design and implementation of a new translation tool that takes \cml code annotated with \gospel specifications and produces a corresponding \whyml representation suitable 
for verification in \whythree. This tool encompasses a complete pipeline including a lexer, parser, AST, Gospel AST, and a printer that maps both \cml constructs and \gospel annotations into their 
\whyml equivalents, giving the already existing \cml library of code a new way to easily verify that code with only some handwritten specifications.

The third and main contribution is a comparative evaluation of the two verification approaches, extraction and translation, using a large and complex case study, that being a verified implementation
of Red-Black trees. This evaluation characterizes the trust boundaries of each approach, identifies the critical points of failure in each pipeline, and discusses the implications for end-to-end 
verification guarantees when combined with \cml's formally verified compiler.

\section{Future Work}

There are several steps required to create a truly robust end-to-end verification pipeline using translation and extraction mechanisms. A portion of our work focused on revising and modernizing
the extraction mechanism found in \whythree's for \cml. However, as previously mentioned, we have identified some limitations with our solution, which we would like to revisit as future work. In
particular, we want to introduce a static verification tool that automatically determines whether a \whyml program operates within the supported subset prior to extraction. This would remove
the current reliance on the user to manually ensure that the \whyml program complies with that subset.

A more ambitious direction would be to prove the unverified components that constitute the critical points of failure in each pipeline. For the extraction approach this would 
mean formally verifying the correctness of the \cml extraction driver, establishing that it faithfully preserves the semantics of the \whyml source in the generated \cml code. For the translation 
approach it would mean formally verifying the translation tool itself, ensuring that the \whyml representation it produces is a faithful model of the original \cml program and its \gospel 
specifications. Addressing these two gaps would significantly strengthen the end-to-end verification guarantees of both pipelines and bring them closer to a fully trustworthy verification 
chain from source code to machine code.